Windowed least-squares model reduction for dynamical systems

### 2019-10-06 main.tex
update

\KTC{HERE}
Despite its time-instantaneous optimality property 

@@

is an alternative projection method that exhibits several advantages over Galerkin

@@

Rather than minimize the (time-continuous) FOM ODE residual as in Galerkin
projection, at each time step, LSPG minimizes
the FOM O$\Delta$E residual, i.e., the residual arising after applying time discretization to the FOM ODE.  

@@

To begin, and without loss of generality, we introduce a uniform time
grid characterized by the time step $\Delta t$ and time instances
$t^n = n\Delta
t$, $n=1,\ldots,N_t$.
Applying a linear multistep method to discretize the FOM ODE \eqref{eq:FOM}
with this time grid
yields the FOM O$\Delta$E, which computes the sequence of FOM O$\Delta$E
solutions
$\stateFOMDiscreteArg{n}( \approx \stateFOM(t^n))$, $n=1,\ldots,N_t$
as the implicit solution to the system of algebraic equations
\begin{equation}\label{eq:lms}
\residLMSArg{n} (\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n}) = \bz, 
\end{equation}
where  $k^n$ denotes the number of steps employed by the scheme at the $n$th
time instance and 
$\residLMSArg{n}$ denotes the FOM O$\Delta$E residual, which is defined as
\begin{align*}
\residLMSArg{n} &: (\stateyDiscreteArgnt{n};\stateyDiscreteArgnt{n-1},\ldots,\stateyDiscreteArgnt{n-k^n}) \mapsto  \frac{1}{\Delta t} \sum_{j=0}^{k^n} \alpha^n_j \stateyDiscreteArgnt{n-j} -  \sum_{j=0}^{k^n} \beta^n_j \velocity(\stateyDiscreteArgnt{n-j},t^{n-j}),
\\
&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim},
\end{align*} 
Here, $\alpha^n_j$, $\beta^n_j$, $j=0,\ldots,k^n$ are coefficients
that define the linear multistep method at the $n$th time instance.
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\residLMS \frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%the algebraic FOM O$\Delta$E obtained after an implicit Euler temporal discretization is given by,
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance.

@@

Substituting the approximation~\eqref{eq:stapprox1} into the FOM ODE~\eqref{eq:FOM},
\begin{equation}\label{eq:st_fom}
\frac{d }{dt} \stbasisArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate + \stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM)(0), \qquad t \in [0,T],
\end{equation}
where $\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 
The space--time Galerkin method is noted to additionally be a possibility for model reduction. As the space--time Galerkin method does not associate with any residual minimization 
principle, however, we do not outline it here.
\subsubsection{Space--Time LSPG projection}  
As in LSPG, Space--Time LSPG (ST-LSPG)~\cite{choi_stlspg} minimizes the fully discrete residual, with the key difference being that ST-LSPG minimizes the residual over the entire time domain. For brevity, ST-LSPG is outlined only for linear multistep methods.
%As ST-LSPG acts on the discrete solution, we first note that the discrete FOM solution lives in $\RR{N} \otimes \RR{N_t}$. ST-LSPG then seeks solutions in the trial space,
%$$\stspaceDiscrete \defeq \spaceTrialSpace \otimes \timeTrialSpaceDiscrete \subseteq \RR{N} \otimes \RR{N_t},$$
%where $\timeTrialSpaceDiscrete$ is the discrete temporal trial space. It is noted that this trial space is a specific instance of the trial space~\ref{eq:sttrialspace_def} 
%As before, we define a discretization of the continuous time-domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time-step $\Delta t = T/N_t$. We define now the space--time discrete residual for the implicit Euler method as,
We first define the space--time discrete residual for linear multistep methods,
\begin{align*}
\residST_{} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \\ 
%\frac{\stateyDiscreteArg{N_t-1} - \stateFOMDiscreteArg{N_t-2}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t-1}) \\
\vdots \\
%\frac{\stateyDiscreteArg{1} - \stateFOMIC }{\Delta t} -  \velocity(\stateyDiscreteArg{1}) \\
\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
\end{bmatrix}  \\
& \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
For convenience we define the same residual with the generalized coordinates as the inputs,
\begin{align*}
\genresidST_{\text{}} & \vcentcolon (\stgenstatey ;\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg) \\ 
\vdots \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^1}\stgenstatey + \stateIntercept; \stateFOMIC \bigg) \\ 
\end{bmatrix}  \\
& \vcentcolon \RR{\stdim} \times \RR{N} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
ST-LSPG is defined as,
\begin{equation*}
\stgenstate = \underset{ \stgenstatey }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{}}(\stgenstatey;\stateFOMIC) ||_2^2, 
\end{equation*}
where $\lspgWeightingSTArg{\cdot} \in \RR{n_{st} \times N N_t}$, with $\stdim \le n_{st} \le N N_t$, is a space--time weighting matrix.

>>>>

Despite its time-instantaneous residual-minimization optimality property 

@@

was developed
as an alternative projection method that exhibits several advantages over Galerkin

@@

Rather than minimize the (time-continuous) FOM ODE residual at a time instance as in Galerkin
projection, LSPG minimizes
the (time-discrete) FOM O$\Delta$E residual 
(i.e., the residual arising after applying time discretization to the FOM ODE)
over a time step.

@@

Without loss of generality, we introduce a uniform time
grid characterized by time step $\Delta t$ and time instances
$t^n = n\Delta
t$, $n=0,\ldots,N_t$.
Applying a linear multistep method to discretize the FOM ODE \eqref{eq:FOM}
with this time grid
yields the FOM O$\Delta$E, which computes the sequence of discrete
solutions
$\stateFOMDiscreteArg{n}( \approx \stateFOM(t^n))$, $n=1,\ldots,N_t$
as the implicit solution to the system of algebraic equations
\begin{equation}\label{eq:lms}
\residLMSArg{n}
	(\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n})
	= \bz,\qquad n=1,\ldots,N_t,
\end{equation}
where  $k^n$ denotes the number of steps employed by the scheme at the $n$th
time instance and 
$\residLMSArg{n}$ denotes the FOM O$\Delta$E residual defined as
\begin{align*}
\residLMSArg{n} &: (\stateyDiscreteArgnt{n};\stateyDiscreteArgnt{n-1},\ldots,\stateyDiscreteArgnt{n-k^n}) \mapsto  \frac{1}{\Delta t} \sum_{j=0}^{k^n} \alpha^n_j \stateyDiscreteArgnt{n-j} -  \sum_{j=0}^{k^n} \beta^n_j \velocity(\stateyDiscreteArgnt{n-j},t^{n-j}),
\\
&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim},
\end{align*} 
Here, $\alpha^n_j,\beta^n_j\in\RR{}$, $j=0,\ldots,k^n$ are coefficients
that define the linear multistep method at the $n$th time instance.
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\residLMS \frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%the algebraic FOM O$\Delta$E obtained after an implicit Euler temporal discretization is given by,
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance.

@@

Substituting the approximation~\eqref{eq:stapprox1} into the FOM ODE~\eqref{eq:FOM},
\begin{equation}\label{eq:st_fom}
\stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate + \stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM)(0), \qquad t \in [0,T],
\end{equation}
where $\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 
The space--time Galerkin method is noted to additionally be a possibility for model reduction. As the space--time Galerkin method does not associate with any residual minimization 
principle, however, we do not outline it here.
\subsubsection{Space--Time LSPG projection}  
As in LSPG, Space--Time LSPG (ST-LSPG)~\cite{choi_stlspg} minimizes the fully discrete residual, with the key difference being that ST-LSPG minimizes the residual over the entire time domain. For brevity, ST-LSPG is outlined only for linear multistep methods.
%As ST-LSPG acts on the discrete solution, we first note that the discrete FOM solution lives in $\RR{N} \otimes \RR{N_t}$. ST-LSPG then seeks solutions in the trial space,
%$$\stspaceDiscrete \defeq \spaceTrialSpace \otimes \timeTrialSpaceDiscrete \subseteq \RR{N} \otimes \RR{N_t},$$
%where $\timeTrialSpaceDiscrete$ is the discrete temporal trial space. It is noted that this trial space is a specific instance of the trial space~\ref{eq:sttrialspace_def} 
%As before, we define a discretization of the continuous time-domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time-step $\Delta t = T/N_t$. We define now the space--time discrete residual for the implicit Euler method as,
We first define the space--time discrete residual for linear multistep methods,
\begin{align*}
\residST_{} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \\ 
%\frac{\stateyDiscreteArg{N_t-1} - \stateFOMDiscreteArg{N_t-2}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t-1}) \\
\vdots \\
%\frac{\stateyDiscreteArg{1} - \stateFOMIC }{\Delta t} -  \velocity(\stateyDiscreteArg{1}) \\
\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
\end{bmatrix}  \\
& \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
For convenience we define the same residual with the generalized coordinates as the inputs,
\begin{align*}
\genresidST_{\text{}} & \vcentcolon (\stgenstatey ;\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg) \\ 
\vdots \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^1}\stgenstatey + \stateIntercept; \stateFOMIC \bigg) \\ 
\end{bmatrix}  \\
& \vcentcolon \RR{\stdim} \times \RR{N} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
ST-LSPG is defined as,
\begin{equation*}
\stgenstate = \underset{ \stgenstatey }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{}}(\stgenstatey;\stateFOMIC) ||_2^2, 
\end{equation*}
where $\lspgWeightingSTArg{\cdot} \in \RR{n_{st} \times N N_t}$, with $\stdim \le n_{st} \le N N_t$, is a space--time weighting matrix.
